Global patterns of forest autotrophic carbon fluxes

### 2020-01-20 manuscript/manuscript-draft.tex
Table 2

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Global patterns of forest autotrophic carbon fluxes},
            pdfauthor={Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene Muller-Landau, Kristina J. Anderson-Teixeira},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Global patterns of forest autotrophic carbon fluxes}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben
Bond-Lamberty, Helene Muller-Landau, Kristina J. Anderson-Teixeira}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

@@

{[}\textbf{very rough start:}{]} \ldots{}.. Here, we drew upon \#
records from the Global Forest Carbon Database (ForC), representing all
major forest types and the nine most significant forest autotrophic
carbon flux (FACF) variables, to comprehensively explore how C cycling
in mature, undisturbed forests varies with latitude and climate on a
global scale. We show that, across all FACF variables analyzed, C
cycling decreases linearly with latitude --a finding that confirms
multiple previous studies but contradicts the idea that productivity of
temperate forests rivals that of tropical forests. The FACF variables
increase in proportion to one another, with no differences in allocation
detectable at this global scale, but differed in that latitude explained
a lower proportion of variation among subsidiary fluxes (in particular,
woody aboveground productivity and belowground net primary productivity,
\(BNPP\)). Climate explained a significant proportion (\#-\#\%) of
variation in all C fluxes analyzed (less for subsidiary fluxes), with
temperature variables in general and mean annual temperature (\(MAT\))
in particular being the best predictors of FACF on this global scale.
While other climate variables (\emph{e.g.}, displayed significant
correlation with FACF, none of them improved significantly over the
explanatory power \(MAT\). The effects of temperature were modified by
moisture availability, with reduced FACF under hot and dry conditions
and sometimes under very high precipitaiton (especially for \(BNPP\)).
FACF declined with temperature seasonality, but growing season length
doesn't improve upon MAT as a predictor. Within the growing season, the
influence of climate on C cycling is smaller but still significant for a
number of carbon fluxes. These findings clarify the big picture of how
FACF varies with latitude and climate on a global scale. As we enter a
period of accelerating climatic change, understanding of the fundamental
climatic controls on FACF sets a foundation for understanding patterns
of change.

@@

\textbf{Forests play a critical role in regulating atmospheric CO\_2 and
climate} (\textbf{Bonan}). The total amount of CO\_2 cycling through
Earth's forests each year (\emph{i.e.}, total gross primary production,
\(GPP\)) is more than five times anthropogenic fossil fuel emissions,
and net sequestration (\textasciitilde{}2.4 GtCyr\textsuperscript{-1})
offsetting roughly 30\% of these emissions \citep{pan_large_2011}. As
atmospheric carbon dioxide levels continue to rise, with consequences
for global climate, there is increasing recognition that proper
protection and management of forest resources will have an important
role to play in mitigating climate change (** REFS- e.g., IPCC 1.5C and
land reports**). Carbon fixation, allocation, and metabolism by trees
set the basis for energy and material flows in forest ecosystems and
define their interactions with Earth's changing climate.

\textbf{Forests' autotrophic carbon flux (FACF, bbl:define) influences
all organic matter stocks in ecosystems and is linked to cycling of
energy, water, and nutrients} (\textbf{REFS; \emph{Piao et al. 2010.
Forest annual carbon cost: a global‐scale analysis of autotrophic
respiration. Ecology 91(3), 652-661Ƴ} }). Forest productivity sets the
energy ultimately available to heterotrophic organisms (including
microbes), in turn influencing their abundance (\textbf{REFS}) and
possibly diversity
(\href{https://www.annualreviews.org/doi/abs/10.1146/annurev.ecolsys.30.1.257}{Waide
et al. 2009}\textbf{REFS})--but probably not individual metabolic rate
(\textbf{Anderson \& Jetz 2005, but check more recent literature}).
However, our ability to draw general macroscopic conclusions regarding
global variation in multiple productivity variables with respect to
climate has been limited in that these analyses often mix forests that
vary in stand age, disturbance history, and/or management status; do not
always sufficiently parse related variables (\emph{e.g.}, combining net
primary productivity records with and without belowground components);
and typically consider only one or a few variables at a time.

\textbf{The productivity of forests varies with latitude}, showing a
general trend of decreasing productivity with latitude
\citep{beer_terrestrial_2010, jung_global_2011}. Studies agree that
productivity is lowest in the boreal regions, and increases into the
temperate regions
\citep{luyssaert_co_2007, huston_global_2009, beer_terrestrial_2010, jung_global_2011}.
However, evidence is inconclusive on whether productivity continues to
increase into the tropics, or whether it plateaus in temperate regions.
Evidence for this is further complicated by the fact that different
studies use different measures of productivity to explore these
relationships. For example, modelling of global terrestrial ecosystem
gross primary productivity (GPP) through upscaling and calibration of
eddy flux measurements indicates that GPP peaks in tropical forests
\citep{beer_terrestrial_2010, jung_global_2011};\href{https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.14729}{Bagdley
et al. 2019}). This is corroborated by analysis of site-level GPP
measurements, which appear to reach their highest levels in tropical
forests \citep{luyssaert_co_2007}. In contrast, there is evidence that
the highest values of net primary productivity (NPP) may be found in
temperate forests \citep{luyssaert_co_2007, huston_global_2009},
although other studies find NPP is highest in the tropics, showing a
decrease with latitude \citep{simova_enigma_2017}. Other studies have
chosen to focus exclusively on above-ground net primary productivity
(ANPP), finding evidence of a weak negative relationship between ANPP
and latitude \citep{huston_global_2009, gillman_latitude_2015}.

\textbf{Climate is a significant driver of productivity across broad
spatial scales} \citep{cleveland_relationships_2011}. The majority of
studies have focused on exploring the relationships between productivity
and mean annual temperature (MAT) and mean annual precipitation (MAP),
as the most commonly reported site-level climate variables. These
variables have the advantage that they describe broad trends in
temperature and water availability, and therefore capture a lot of
global-scale variation in climate. There is strong evidence that both
MAT and MAP show significant positive relationships with productivity
\citep{chu_does_2016}. However, as with latitude, the shape of those
relationships is not always clear, and, again, is complicated by the use
of different measures of productivity across studies. Various measures
of primary productivity saturate at high levels of MAP, though the
saturation points identified vary from 1500mm \citep{luyssaert_co_2007}
up to 2445mm MAP \citep{schuur_productivity_2003}. Studies of the
influence of MAT on productivity are less conclusive. Luyssaert et al.
\citeyearpar{luyssaert_co_2007} examined GPP and NPP and found that,
while GPP increases linearly with MAT, NPP saturates at around
10\(^\circ\)C MAT. In contrast, Larjavaara and Muller-Landau
\citeyearpar{larjavaara_temperature_2012}, find that increases in GPP
saturate at approximately 25\(^\circ\)C MAT, while Schuur
\citeyearpar{schuur_productivity_2003} shows that NPP increases linearly
with temperature. bbl:maybe remove following sentences The influence of
these climate variables on productivity is further complicated by the
possibility of interactive effects occurring between them. Taylor et al.
\citeyearpar{taylor_temperature_2017} showed that increased MAP had a
negative effect at low MAT, but a positive effect at high levels of MAT,
and vice versa, suggesting that both temperature and water availability
are important in explaining productivity.

\textbf{Primary productivity can be influenced by many other factors as
well}, which often act across a range of scales, and may show
interactive effects with each other
\citep{cleveland_relationships_2011}. On a local scale, stand age
\citep{litton_carbon_2007, gillman_latitude_2015}, biodiversity (Liang
et al. 2019. Positive biodiversity-productivity relationship predominant
in global forests. Science 354, Issue 6309, aaf8957), management
\citep{simova_enigma_2017}; nutrient availability
\citep{aragao_above-_2009}; and altitude
\citep{girardin_net_2010, malhi_variation_2017} all impact forest
productivity. On a global scale, we expect that productivity is most
strongly influenced by broad climatic gradients. There is evidence that
productivity also responds to variables such as cloud cover
\citep{taylor_temperature_2017}, solar radiation
\citep{fyllas_solar_2017}, and potential evapotranspiration
\citep{kerkhoff_plant_2005} in potentially significant ways.
Furthermore, MAT and MAP are very coarse measures of climate, and so
fail to capture much variation in climate on an intra-annual scale,
including the effects of factors such as growing season length, number
of frost-free days, temperature seasonality, and dry season length. Some
studies have suggested that the apparently strong relationship between
MAT and productivity is actually a factor of the correlation between MAT
and growing season length
\citep{kerkhoff_plant_2005, malhi_productivity_2012, michaletz_convergence_2014, michaletz_drivers_2018}.
Kerkhoff et al. \citeyearpar{kerkhoff_plant_2005} and Michaletz et al.
\citeyearpar{michaletz_convergence_2014} find that, within the growing
season, there is no significant relationship between productivity and
MAT, indicating that the effect of temperature is due to increased
length of growing season, rather than an inherent influence of
temperature on productivity.

\textbf{The recent development of a global forest carbon database
synthesizing multiple variables and including records of stand history
(ForC;
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018})
opens up the possibility for a standardized analysis of global scale
variation} in multiple components of forest productivity and the
principle climatic drivers of these patterns. In order to approach these
broad and complex issues, we simplify the major gaps in our knowledge to
five broad hypotheses and corresponding specific predictions (Table 1).
First, we ask how forest autotrophic carbon fluxes (FACF) vary with
latitude. We then test how these fluxes relate to MAT and MAP, and
additionally how they respond to other, less well studied, climate
variables. We finally consider the relationship between FACF and
seasonality, considering the role of seasonality in explaining variation
in carbon fluxes, and the influence of climate on FACF per month of the
growing season. We use a comprehensive global database of forest carbon
fluxes to explore the above questions for nine carbon fluxes, allowing
for an in-depth exploration of the effect of climate on global
productivity.

@@

\hspace{1em}1.1. FACFs decrease linearly with latitude (L-) & Luyssaert et al. (2007) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 2-3\\
\hspace{1em}1.1.alt. FACFs are similar in tropical and temperate forests, but lower in boreal regions (CD-) & Gillman et al. (2015); Simova and Storch (2017) & no & no & no & no & no & no & no & no & no & no & Fig. 2-3\\

@@

\hspace{1em}2.1. FACFs increase linearly with MAT (L+) & Schuur (2003) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 5\\
\hspace{1em}2.2. FACFs increase with precipitation but saturate or decrease at very high levels (CD+) & Luyssaert et al. (2007); Schuur (2003) & yes & yes & yes & yes & yes & yes & n.s. & L+ & L+ & yes & Fig. 5\\
\hspace{1em}2.3. There is a postive interaction between temperature and precipitation (I+) & Taylor et al. (2016) & (yes) & yes & L+ & yes & L- & L+ & yes & L- & yes & yes & Fig. 4\\

@@

\hspace{1em}3.1. FACFs increases with PET, but saturates or decreases at high levels (CD+) &  & (yes) & yes & yes & yes & yes & L+ & yes & L+ & yes & L+ & Fig. 5\\
\hspace{1em}3.2. FACFs increase with vapour pressure deficit, but saturate or decrease at high levels (CD+) &  & (yes) & yes & yes & yes & yes & yes & yes & L+ & yes & yes & Fig. 5\\

@@

\hspace{1em}4.1. FACFs decrease linearly with temperature seasonality (L-) &  & yes & CU+ & CU+ & CU+ & yes & yes & yes & yes & CU+ & CU+ & Fig. 5\\

@@

\hspace{1em}4.3. FACFs increase linearly with growing season length (L+) & Malhi (2012); Michaletz et al. (2014); Chu et al. (2016) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 5\\

@@

Analyses were conducted on data contained in the open-access ForC
database
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018}.
This database contains records of field-based measurements of forest
carbon stocks and annual fluxes, compiled from original publications and
existing data compilations and databases. Associated data, such as stand
age, measurement methodologies, and disturbance history, are also
included. The database was significantly expanded since the publication
of \citep{anderson-teixeira_forc:_2018} through integration with the
Global Database of Soil Respiration Database
\citep{bond-lamberty_global_2010}. Additional targeted literature
searches were conducted to identify any further available data on
primary productivity, with particular focus on mature forests in
temperate and boreal regions. ForC currently contains 29730 records from
4979 plots, representing 20 distinct ecozones across all forested
biogeographic and climate zones. We used ForC v3.0, archived on Zenodo
with DOI 10.5281/zenodo.3403855.

\begin{figure}[H]
\includegraphics[width=1\linewidth]{distribution_all_variables} \caption{Map showing all data used in the analysis, coded by variable}\label{fig:unnamed-chunk-5}
\end{figure}

\emph{Data selection.} Over 50 variables of forest carbon stocks and
annual fluxes are represented in the ForC database; this analysis
focussed on measures of primary productivity. Table 1 contains details
of the variables selected for analysis.

@@

\caption{Definitions and sample sizes of variables used in analysis.
Geographic areas group geographically proximate sites, defined using a
hierarchical cluster analysis on the distance matrix of the sites, and a
cutoff of 25km.}\tabularnewline

@@

Annual gross primary production; annual uptake of carbon dioxide by an
ecosystem\strut

@@

Flux partitioning of eddy covariance\strut

@@

Annual net primary production; the component of GPP that is stored in
plant tissue; \(GPP - R_{auto}\)\strut

@@

Foliage, branch, stem, coarse root and fine root\strut

@@

Direct measurement of annual increments of components\strut

@@

Aboveground net primary production\strut

@@

Foliage, stem, and optionally branch\strut

@@

Direct measurement of annual increments of components\strut

@@

@@

Net primary production of foliage\strut

@@

Foliage\strut

@@

Direct measurement of litterfall, correcting for changes in leaf biomass
when measured\strut

@@

Net primary production of woody stems\strut
Woody stems\strut
Direct measurement of stem growth increment\strut

@@

Belowground net primary production\strut

@@

Coarse and fine roots\strut

@@

Direct measurement of one or more of: fine root turnover, soil cores,
root ingrowth cores, minirhizotrons; indirect estimates of coarse roots
using allometries based on aboveground stem increment measures\strut

@@

Net primary production of fine roots\strut

@@

Fine roots\strut

@@

Direct measurement of one or more of: minirhizotrons, fine root
turnover, soil cores, root ingrowth cores\strut

@@

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Global patterns of forest autotrophic carbon fluxes},
            pdfauthor={Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene C. Muller-Landau, Kristina J. Anderson-Teixeira},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Global patterns of forest autotrophic carbon fluxes}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben
Bond-Lamberty, Helene C. Muller-Landau, Kristina J. Anderson-Teixeira}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

@@

{[}\textbf{very rough start:}{]} \ldots{}.. Here, we drew upon \#
records from the Global Forest Carbon Database (ForC), representing all
major forest types and the nine most significant forest autotrophic
carbon flux (FACF) variables, to comprehensively explore how C cycling
in mature, undisturbed forests varies with latitude and climate on a
global scale. We show that, across all FACF variables analyzed, C
cycling decreases linearly with latitude --a finding that confirms
multiple previous studies and contradicts the idea that productivity of
temperate forests rivals that of tropical forests. The FACF variables
increase in proportion to one another, with no differences in allocation
detectable at this global scale, but differed in that latitude explained
a lower proportion of variation among subsidiary fluxes (in particular,
woody aboveground productivity and belowground net primary productivity,
\(BNPP\)). Climate explained a significant proportion (\#-\#\%) of
variation in all C fluxes analyzed (less for subsidiary fluxes), with
temperature variables in general and mean annual temperature (\(MAT\))
in particular being the best predictors of FACF on this global scale.
While other climate variables (\emph{e.g.}, \textbf{XX}) displayed
significant correlation with FACF, none of them had significantly better
explanatory power than \(MAT\). The effects of temperature were modified
by moisture availability, with reduced FACF under hot and dry conditions
and sometimes under very high precipitaiton (especially for \(BNPP\)).
FACF declined with temperature seasonality, but growing season length
doesn't improve upon MAT as a predictor. Within the growing season, the
influence of climate on C cycling is smaller but still significant for a
number of carbon fluxes. These findings clarify the big picture of how
FACF varies with latitude and climate on a global scale. As we enter a
period of accelerating climatic change, understanding of the fundamental
climatic controls on FACF sets a foundation for understanding patterns
of change.

@@

(\textbf{revisit these first two paragraphs.})

\textbf{Forests play a critical role in regulating atmospheric
CO\textsubscript{2} and climate} (\textbf{Bonan}). Approximately
\textbf{\#\#} Gt of CO\textsubscript{2} cycle through Earth's forests
each year \textbf{REF-KAT}), and their net C sequestration
(\textasciitilde{}2.4 GtCyr\textsuperscript{-1}) offsets roughly 30\% of
anthropogenic fossil fuel emissions \citep{pan_large_2011}. As
atmospheric carbon dioxide levels continue to rise, driving climate
change, there is increasing recognition that protection and management
of forest resources will play an important role in mitigating climate
change (** REFS- e.g., IPCC 1.5C and land reports**). Carbon fixation,
allocation, and metabolism by trees set the basis for energy and
material flows in forest ecosystems and define their interactions with
Earth's changing climate.

\textbf{Forests' autotrophic carbon flux (FACF, bbl:define) influences
all organic matter stocks in ecosystems and is linked to cycling of
energy, water, and nutrients} (\textbf{REFS; \emph{Piao et al. 2010.
Forest annual carbon cost: a global‐scale analysis of autotrophic
respiration. Ecology 91(3), 652-661} }). Forest productivity sets the
energy ultimately available to heterotrophic organisms (including
microbes), in turn influencing their abundance (\textbf{REFS}) and
possibly diversity
(\href{https://www.annualreviews.org/doi/abs/10.1146/annurev.ecolsys.30.1.257}{Waide
et al. 2009}\textbf{REFS}). However, our ability to draw general
macroscopic conclusions regarding global variation in multiple
productivity variables with respect to climate has been limited in that
these analyses often mix forests that vary in stand age, disturbance
history, and/or management status; do not always sufficiently parse
related variables (\emph{e.g.}, combining net primary productivity
records with and without belowground components); and typically consider
only one or a few variables at a time.

\textbf{The productivity of forests varies with latitude}, showing a
general trend of decreasing productivity with latitude
\citep{beer_terrestrial_2010, jung_global_2011}. Studies agree that
productivity is lowest in the boreal regions, and increases into the
temperate regions
\citep{luyssaert_co_2007, huston_global_2009, beer_terrestrial_2010, jung_global_2011}.
However, evidence is inconclusive on whether productivity continues to
increase into the tropics, or whether it plateaus in temperate regions.
Evidence for this is further complicated by the fact that different
studies use different measures of productivity to explore these
relationships. For example, modelling of global terrestrial ecosystem
gross primary productivity (GPP) through upscaling and calibration of
eddy flux measurements indicates that GPP peaks in tropical forests
\citep{beer_terrestrial_2010, jung_global_2011};\href{https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.14729}{Bagdley
et al. 2019}). This is corroborated by analysis of site-level GPP
measurements, which appear to reach their highest levels in tropical
forests \citep{luyssaert_co_2007}. \{\textbf{condense the previous two
sentences into one-- both primarily eddy flux}\} In contrast, some
studies suggest that the highest values of net primary productivity
(NPP) may be found in temperate forests
\citep{luyssaert_co_2007, huston_global_2009}, while others find NPP
highest in the tropics and decreasing with latitude
\citep{simova_enigma_2017}. Other studies have chosen to focus
exclusively on above-ground net primary productivity (ANPP), finding
evidence of a weak negative relationship between ANPP and latitude
\citep{huston_global_2009, gillman_latitude_2015}.

\textbf{Climate is a significant driver of productivity across broad
spatial scales} \citep{cleveland_relationships_2011}. The majority of
studies have focused on exploring the relationships of productivity to
mean annual temperature (MAT) and mean annual precipitation (MAP), as
the most commonly reported site-level climate variables. These variables
have the advantage that they describe broad trends in temperature and
water availability, and therefore capture a lot of global-scale
variation in climate. There is strong evidence that both MAT and MAP
show significant positive relationships with productivity
\citep{chu_does_2016}. However, as with latitude, the shape of those
relationships is not always clear, and, again, is complicated by the use
of different measures of productivity across studies. Various measures
of primary productivity saturate at high levels of MAP, though the
saturation points identified vary from 1500mm \citep{luyssaert_co_2007}
up to 2445mm MAP \citep{schuur_productivity_2003}. Studies of the
influence of MAT on productivity are less conclusive. Luyssaert et al.
\citeyearpar{luyssaert_co_2007} examined GPP and NPP and found that,
while GPP increases linearly with MAT, NPP saturates at around
10\(^\circ\)C MAT. In contrast, Larjavaara and Muller-Landau
\citeyearpar{larjavaara_temperature_2012}, find that increases in GPP
saturate at approximately 25\(^\circ\)C MAT, while Schuur
\citeyearpar{schuur_productivity_2003} shows that NPP increases linearly
with temperature. bbl:maybe remove following sentences The influence of
these climate variables on productivity is further complicated by the
possibility of interactive effects occurring between them. Taylor et al.
\citeyearpar{taylor_temperature_2017} showed that increased MAP had a
negative effect at low MAT, but a positive effect at high levels of MAT,
and \emph{vice versa} \{\textbf{Helene found this confusing}\},
suggesting that both temperature and water availability are important in
explaining productivity.

\textbf{Primary productivity can be influenced by many other factors as
well}, which often act across a range of scales, and may show
interactive effects with each other
\citep{cleveland_relationships_2011}. On a local scale, stand age
\citep{litton_carbon_2007, gillman_latitude_2015}, biodiversity (Liang
et al. 2019. Positive biodiversity-productivity relationship predominant
in global forests. Science 354, Issue 6309, aaf8957), management
\citep{simova_enigma_2017}; nutrient availability
\citep{aragao_above-_2009}; and altitude
\citep{girardin_net_2010, malhi_variation_2017} all impact forest
productivity. On a global scale, we expect that productivity is most
strongly influenced by broad climatic gradients. There is evidence that
productivity also responds to variables such as cloud cover
\citep{taylor_temperature_2017}, solar radiation
\citep{fyllas_solar_2017}, and potential evapotranspiration
\citep{kerkhoff_plant_2005} in potentially significant ways.
Furthermore, MAT and MAP are very coarse measures of climate, and so
fail to capture much variation in climate on an intra-annual scale,
including the effects of factors such as growing season length, number
of frost-free days, temperature seasonality, and dry season length. Some
studies have suggested that the apparently strong relationship between
MAT and productivity is actually a consequence of the correlation
between MAT and growing season length
\citep{kerkhoff_plant_2005, malhi_productivity_2012, michaletz_convergence_2014, michaletz_drivers_2018}.
Kerkhoff et al. \citeyearpar{kerkhoff_plant_2005} and Michaletz et al.
\citeyearpar{michaletz_convergence_2014} find that, within the growing
season, there is no significant relationship between productivity and
MAT, suggesting that the effect of temperature is due to increased
length of growing season, rather than an inherent influence of
temperature on productivity.

\textbf{The recent development of a global forest carbon database
synthesizing multiple variables and including records of stand history
(ForC;
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018})
opens up the possibility for a standardized analysis of global scale
variation} in multiple components of forest productivity and the
principle climatic drivers of these patterns. In order to approach these
broad and complex issues, we simplify the major gaps in our knowledge to
five broad hypotheses and corresponding specific predictions (Table 1).
** see
\href{https://github.com/forc-db/Global_Productivity/issues/48}{issue
\#48}** First, we ask how forest autotrophic carbon fluxes (FACF) vary
with latitude. We then test how these fluxes relate to MAT and MAP, and
additionally how they respond to other, less well studied, climate
variables. Finally, we consider the relationship between FACF and
seasonality, considering the role of seasonality in explaining variation
in carbon fluxes, and the influence of climate on FACF standardized by
growing season length. We use a comprehensive global database of forest
carbon fluxes to address the above questions for nine carbon fluxes,
allowing for an in-depth exploration of the effect of climate on global
productivity.

@@

\hspace{1em}1.1. FACFs decrease linearly with latitude (L-) & Luyssaert et al. (2007) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 2\\
\hspace{1em}1.1.alt. FACFs are similar in tropical and temperate forests, but lower in boreal regions (CD-) & Gillman et al. (2015); Simova and Storch (2017) & no & no & no & no & no & no & no & no & no & no & Fig. 2\\

@@

\hspace{1em}2.1. FACFs increase linearly with MAT (L+) & Schuur (2003) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 4\\
\hspace{1em}2.2. FACFs increase with precipitation but saturate or decrease at very high levels (CD+) & Luyssaert et al. (2007); Schuur (2003) & yes & yes & yes & yes & yes & yes & n.s. & L+ & L+ & yes & Fig. 4\\
\hspace{1em}2.3. There is a postive interaction between temperature and precipitation (I+) & Taylor et al. (2016) & (yes) & yes & L+ & yes & L- & L+ & yes & L- & yes & yes & Fig. 3\\

@@

\hspace{1em}3.1. FACFs increases with PET, but saturates or decreases at high levels (CD+) &  & (yes) & yes & yes & yes & yes & L+ & yes & L+ & yes & L+ & Fig. 4\\
\hspace{1em}3.2. FACFs increase with vapour pressure deficit, but saturate or decrease at high levels (CD+) &  & (yes) & yes & yes & yes & yes & yes & yes & L+ & yes & yes & Fig. 4\\

@@

\hspace{1em}4.1. FACFs decrease linearly with temperature seasonality (L-) &  & yes & CU+ & CU+ & CU+ & yes & yes & yes & yes & CU+ & CU+ & Fig. 4\\

@@

\hspace{1em}4.3. FACFs increase linearly with growing season length (L+) & Malhi (2012); Michaletz et al. (2014); Chu et al. (2016) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 4\\

@@

Analyses were conducted on data contained in the open-access ForC
database
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018}.
This database contains records of field-based measurements of forest
carbon stocks and annual fluxes, compiled from original publications and
existing data compilations and databases. Associated data, such as stand
age, measurement methodologies, and disturbance history, are also
included. The database was significantly expanded since the publication
of \citep{anderson-teixeira_forc:_2018} through integration with the
Global Soil Respiration Database \citep{bond-lamberty_global_2010}.
Additional targeted literature searches were conducted to identify any
further available data on primary productivity, with particular focus on
mature forests in temperate and boreal regions. ForC currently contains
29730 records from 4979 plots, representing 20 distinct ecozones across
all forested biogeographic and climate zones. We used ForC v3.0,
archived on Zenodo with DOI 10.5281/zenodo.3403855.

\begin{figure}[H]
\includegraphics[width=1\linewidth]{distribution_all_variables_cropped} \caption{Map showing all data used in the analysis, coded by variable. Variables are plotted individually in Fig. S1. }\label{fig:unnamed-chunk-5}
\end{figure}

\emph{Data selection.} Over 50 variables of forest carbon stocks and
annual fluxes are represented in the ForC database; this analysis
focused on autotrophic carbon fluxes. Focal variables of this analysis
are listed in Table 2.

A subset of the ForC database was generated for the purposes of this
analysis, in order to control for data quality and remove biasing
factors. Since management can alter observed patterns of primary
productivity \citep{simova_enigma_2017}, sites were excluded from
analysis if they were managed, defined as plots that were planted,
managed as plantations, irrigated, fertilised or including the term
``managed'' in their site description. Sites that had experienced
significant disturbance were also excluded. Disturbances that justified
site exclusion were major cutting or harvesting, and/or burning,
flooding, drought and storm events with site mortality
\textgreater{}10\% of trees. Grazed sites were retained.\\
There is evidence that stand age influences patterns of primary
productivity and carbon allocation in forest ecosystems, and can
confound relationships between latitude and primary productivity
\citep{de_lucia_forest_2007, gillman_latitude_2015}. To reduce any
biasing effects of stand age, we included only stands of known age
\(\ge\) 100 years and those described by terms such as ``mature'',
``intact'', or ``old-growth''.

\newpage

@@

\caption{Definitions and sample sizes of variables (all annual) used in
analysis. Geographic areas group geographically proximate sites, defined
using a hierarchical cluster analysis on the distance matrix of the
sites, and a cutoff of 25km.}\tabularnewline

@@

Gross Primary Production\strut

@@

flux partitioning of eddy-covariance; \(NPP\)+\(R_{auto}\)\strut

@@

Net Primary Production\strut

@@

stem, foliage, coarse root, fine root, optionally others (e.g., branch,
reproductive, understory)\strut

@@

\(ANPP\) + \(BNPP\) (majority); \(GPP\)-\(R_{auto}\)\strut

@@

Aboveground Net Primary Production\strut

@@

stem, foliage, optionally others (e.g., branch, reproductive,
understory) (\textbf{issues
\href{https://github.com/forc-db/Global_Productivity/issues/52}{52},
\href{https://github.com/forc-db/Global_Productivity/issues/54}{54}})\strut

@@

\(ANPP_{woody-stem}\) + \(ANPP_{foliage}\) (+ others)\strut

@@

Net Primary Production of woody stems\strut
woody stems down to DBH \(\le\) 10cm (no branch turnover)\strut
stem growth measurements scaled to biomass using allometries
(\textbf{with recruitment?})\strut

@@

Net Primary Production of foliage\strut

@@

foliage\strut

@@

litterfall collection (separated into components)\strut

@@

@@

Belowground Net Primary Production\strut

@@

coarse and fine roots\strut

@@

coarse roots estimated indirectly using allometries based on aboveground
stem increment measures ; fine roots as below\strut

@@

Net Primary Production of fine roots\strut

@@

fine roots\strut

@@

measurements combined one or more of the following: minirhizotrons, fine
root turnover, soil cores, root ingrowth cores\strut

@@

Root Respiration\strut
coarse and fine roots\strut
